---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(lubridate)
library(RSocrata)
```
## Get Data
The data is obtained from the city of Chicago website. We will collect the raw data using the exposed API they have made availible. 
```{r}
years_ago <- today() - years(2)
crash_url <- glue::glue("https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if?$where=CRASH_DATE > '{years_ago}'")
crash_raw <- as_tibble(read.socrata(crash_url))
```
We will use skim to get an overview of our data. It is an extremely useful function to examine distributions, and missing values.
```{r}
library(skimr)
skim(crash_raw)
```
Generate our outcome column.  We are going to predict if there was an injury or not in the crash. 
```{r}
crash <- crash_raw %>%
  arrange(desc(crash_date)) %>%
  transmute(
    injuries = if_else(injuries_total > 0, "injuries", "none"),
    crash_date,
    crash_hour,
    report_type = if_else(report_type == "", "UNKNOWN", report_type),
    num_units,
    posted_speed_limit,
    weather_condition,
    lighting_condition,
    roadway_surface_cond,
    first_crash_type,
    trafficway_type,
    prim_contributory_cause,
    latitude, longitude
  ) %>%
  na.omit()

crash
```
How does the number of injuries change week to week? 

```{r}
crash %>%
  # change the crssh_date column with mutate
  mutate(crash_date = floor_date(crash_date, unit = "week")) %>%
  # count the number of crashes
  count(crash_date, injuries) %>%
  # filter out first and last
  filter(
    crash_date != last(crash_date),
    crash_date != first(crash_date)
  ) %>%
  # plot
  ggplot(aes(crash_date, n, color = injuries)) +
  geom_line(size = 1.5, alpha = 0.7) +
  scale_y_continuous(limits = (c(0, NA))) +
  labs(
    x = NULL, y = "Number of traffic crashes per week",
    color = "Injuries?"
  )
```

```{r}
crash %>%
  # mutate to makem date weekly
  mutate(crash_date = floor_date(crash_date, unit = "week")) %>%
  # get count (will be n)
  count(crash_date, injuries) %>%
  filter(
    crash_date != last(crash_date),
    crash_date != first(crash_date)
  ) %>%
  # group by week
  group_by(crash_date) %>%
  # percent of injuries
  mutate(percent_injury = n / sum(n)) %>%
  # ungroup 
  ungroup() %>%
  # we only want ones with unjuries
  filter(injuries == "injuries") %>%
  ggplot(aes(crash_date, percent_injury)) +
  geom_line(size = 1.5, alpha = 0.7, color = "midnightblue") +
  scale_y_continuous(limits = c(0, NA), labels = percent_format()) +
  labs(x = NULL, y = "% of crashes that involve injuries")
```

Does the day of the week affect crashes??
```{r}
crash %>%
  mutate(crash_date = wday(crash_date)) %>%
  count(crash_date,injuries) %>%
  filter(injuries == "injuries") %>%
  ggplot(aes(crash_date,n)) +geom_line()
```

```{r}
crash %>%
  mutate(crash_date = wday(crash_date, label = TRUE)) %>%
  count(crash_date, injuries) %>%
  group_by(injuries) %>%
  mutate(percent = n / sum(n)) %>%
  ungroup() %>%
  ggplot(aes(percent, crash_date, fill = injuries)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_x_continuous(labels = percent_format()) +
  labs(x = "% of crashes", y = NULL, fill = "Injuries?")
```
How do injuries vary with first crash type?

```{r}
crash %>%
  count(first_crash_type, injuries) %>%
  mutate(first_crash_type = fct_reorder(first_crash_type, n)) %>%
  group_by(injuries) %>%
  mutate(percent = n / sum(n)) %>%
  ungroup() %>%
  group_by(first_crash_type) %>%
  filter(sum(n) > 1e4) %>%
  ungroup() %>%
  ggplot(aes(percent, first_crash_type, fill = injuries)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_x_continuous(labels = percent_format()) +
  labs(x = "% of crashes", y = NULL, fill = "Injuries?")
```
Injuries by location
```{r}
crash %>%
  filter(latitude > 0) %>%
  ggplot(aes(longitude, latitude, color = injuries)) +
  geom_point(size = 0.5, alpha = 0.4) +
  labs(color = NULL) +
  scale_color_manual(values = c("deeppink4", "gray80")) +
  coord_fixed()
```
## Building a Model

### Split the data
```{r}
library(tidymodels)

set.seed(42)
crash_split <- initial_split(crash, strata = injuries)
crash_train <- training(crash_split)
crash_test <- testing(crash_split)

set.seed(123)
crash_folds <- vfold_cv(crash_train, strata = injuries)
crash_folds
```
```{r}
library(themis)
library(baguette)

crash_rec <- recipe(injuries ~ ., data = crash_train) %>%
  step_date(crash_date) %>%
  step_rm(crash_date) %>%
  step_other(weather_condition, first_crash_type,
    trafficway_type, prim_contributory_cause,
    other = "OTHER"
  ) %>%
  step_downsample(injuries)

bag_spec <- bag_tree(min_n = 10) %>%
  set_engine("rpart", times = 25) %>%
  set_mode("classification")

crash_wf <- workflow() %>%
  add_recipe(crash_rec) %>%
  add_model(bag_spec)

crash_wf
```
Downsampling (in this context) means training on a disproportionately low subset of the majority class examples.



Let’s fit this model to the cross-validation resamples to understand how well it will perform.

```{r}
doParallel::registerDoParallel()
crash_res <- fit_resamples(
  crash_wf,
  crash_folds,
  control = control_resamples(save_pred = TRUE)
)
```

## Evaluate Model
```{r}
collect_metrics(crash_res)
```

```{r}
crash_fit <- last_fit(crash_wf, crash_split)
collect_metrics(crash_fit)
```
```{r}
## variable importance scores
crash_fit %>% extract_fit_parsnip()
```

```{r}
crash_imp <- crash_fit$.workflow[[1]] %>%
  pull_workflow_fit()

crash_imp$fit$imp %>%
  slice_max(value, n = 10) %>%
  ggplot(aes(value, fct_reorder(term, value))) +
  geom_col(alpha = 0.8, fill = "midnightblue") +
  labs(x = "Variable importance score", y = NULL)
```
```{r}
collect_predictions(crash_fit) %>%
  roc_curve(injuries, .pred_injuries) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(size = 1.5, color = "midnightblue") +
  geom_abline(
    lty = 2, alpha = 0.5,
    color = "gray50",
    size = 1.2
  ) +
  coord_equal()
```
Save Model
```{r}
crash_wf_model <- crash_fit$.workflow[[1]]
```

